### 2015-07-06 narrative/data-reduction.tex
small change

In the case of C. S. Chang described earlier, in commmon with most other
applications scientists with whom we collaborate, information is provided on
which data should be sent over to our SSIO layer in order that the best
possible information is available the subsequent analysis. The decision as how
this should be accomplished generally involves refactoring (re-organizing and
reducing the data).

@@

Ideally, the scientist would perform the entire analysis stage {\em in situ} to
extract the relevant information and, in the unlikely event of this being
possible, would effectively circumvent the large data issue completely.  The
problem of course is that this is unlikely to be possible since, by their
nature, large scale simulations aim to discover new information that is often
hidden in the form of higher order effects amongst the data deluge. In
particular, this means if data thinning or truncation is applied willy nilly
then one is in danger of effectively throwing out the baby with the bath water.
One cannot store the entire data set in an easily accessible source, due to
sheer size, yet one cannot reduce it prior to archival without risking losing
the information one is trying to extract.  Viewed in this way, the problem
would appear intractable. However, the above discussion fails to recognize that
much of the data is redundant in the information theoretic sense.  That is to
say, the amount of information contained in the data is often significantly
less than the amount of data.  The difficulty stems from the fact that one does
not know about this redundancy in advance \emph{without the benefit of 
priori knowledge}. This knowledge can often be provided by the user or can 
be acquired through experience of dealing with differing runs of the same
code.

@@

\paragraph{Proposed research approach:} We will be looking at a multi-throng
approach to data-refactoring. One approach allows the user to specify
information about their data to indicate if the data is spatially varying,
space-and time varying, just time varying. This allows us to store information
in NVRAM on each node (when and where necessary) to later preform compression
in both space and time . We will examine the different techniques for
compression and re-organization along with examine techniques to use an
error-estimator (gradients) to create a multi-resolution view of the data which
can then be sampled, and placed in different quantiles of importance. These
will then be mapped over into the different layers of the storage hierarchy.
We will also investigate technique which can use ``plug-ins'' for detection of
features and store higher resolution data around the feature and lower
resolution values which are further away from the features. This
re-organization must occur without increasing the total amount of storage than
what the users intended. We will also investigate the time it takes to
re-organize the data and use the multi-layer approach to staging to understand
where the data gets reduced and where it gets re-organized. Open questions are:
1) do we move the data over to other nodes to then reduce and re-organize, 2)
do we first reduce the data and then later re-organize this data on different
nodes?  3) do we use different cores on a node to do data-refactoring.
Similarly we will examine the cost of reading the data. It is clear that if we
only read in the high-importance data it should improve the overall time to
move the data from the SSIO layer to the application, but what happens when
users want all of the data? There is an obvious penalty of having to take data
in a hierarchical layout and then place this back to a single-layer view.
Additional research questions which must be addressed: Will some types of data
(PIC data, finite element data, ...) be be conducive  for data-refactoring than
others? How much semantic-domain knowledge will we need to get a dramatic
improvement of the storage system over not doing anything special?

>>>>

In the case of C. S. Chang described earlier, in common with most other
applications scientists with whom we collaborate, information is provided on
which data should be sent over to our SSIO layer in order that the best
possible information is available for the subsequent analysis. The decision as how
this should be accomplished generally involves data refactoring (re-organizing and
reducing the data).

@@

Ideally, the scientists would perform the entire analysis stage {\em in situ} to
extract the relevant information and, in the unlikely event of this being
possible, would effectively circumvent the large data issue completely.  The
problem of course is that this is unlikely to be possible since, by their
nature, large scale simulations aim to discover new information that is often
hidden in the form of higher order effects amongst the data deluge. In
particular, this means if data thinning or truncation is applied willy nilly
then one is in danger of effectively throwing out the baby with the bath water.
One cannot store the entire data set in an easily accessible source, due to
sheer size, yet one cannot reduce it prior to archival without risking losing
the information one is trying to extract.  Viewed in this way, the problem
would appear intractable. However, the above discussion fails to recognize that
much of the data is redundant in the information theoretic sense.  That is to
say, the amount of information contained in the data is often significantly
less than the amount of data.  The difficulty stems from the fact that one does
not know about this redundancy in advance \emph{without the benefit of 
priori knowledge}. This knowledge can often be provided by the user or can 
be acquired through experience of dealing with differing runs of the same
code.

@@

\paragraph{Proposed research approach:} We will be looking at a multi-throng
approach to data-refactoring. One approach allows the user to specify
information about their data to indicate if the data is spatially varying,
space-and time varying, just time varying. This allows us to store information
in NVRAM on each node (when and where necessary) to later preform compression
in both space and time.

We will examine the different techniques for
compression and re-organization along with examine techniques to use an
error-estimator (gradients) to create a multi-resolution view of the data which
can then be sampled, and placed in different quantiles of importance. These
will then be mapped over into the different layers of the storage hierarchy.

We will also investigate technique which can use ``plug-ins'' for detection of
features and store higher resolution data around the feature and lower
resolution values which are further away from the features. This
re-organization must occur without increasing the total amount of storage than
what the users intended.

We will also research the time it takes to
re-organize the data and use the multi-layer approach to staging to understand
where the data gets reduced and where it gets re-organized. Open questions are:
1) do we move the data over to other nodes to then reduce and re-organize, 2)
do we first reduce the data and then later re-organize this data on different
nodes?  3) do we use different cores on a node to do data-refactoring.
Similarly we will examine the cost of reading the data. It is clear that if we
only read in the high-importance data it should improve the overall time to
move the data from the SSIO layer to the application, but what happens when
users want all of the data? There is an obvious penalty of having to take data
in a hierarchical layout and then place this back to a single-layer view.
Additional research questions which must be addressed: Will some types of data
(PIC data, finite element data, ...) be be conducive  for data-refactoring than
others? How much semantic-domain knowledge will we need to get a dramatic
improvement of the storage system over not doing anything special?
